---
title: "12. Looking at Data"
author: "Solutions to Swirl's R Programming Exercises"
date: "07-08-2022"
output: pdf_document
---

Acknowledgements: R Language Concepts and code questions (with minor modifications) are used here from the swirl package. https://www.r-project.org/nosvn/pandoc/swirl.html

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Important note: We don't require to use library(swirl) and swirl() here because we are not going to run the R
script in RStudio Console.

Whenever you're working with a new dataset, the first thing you should do is look at it! What is the format of the
data? What are the dimensions? What are the variable names? How are the variables stored? Are there missing data? Are
there any flaws in the data?

This lesson will teach you how to answer these questions and more using R's built-in functions. We'll be using a
dataset constructed from the United States Department of Agriculture's PLANTS Database
(http://plants.usda.gov/adv_search.html).

```{r}
load("C:/r-basics/Data/plants.rdata")
```
Type ls() to list the variables in your workspace, among which should be plants.

```{r}
ls()
```
Let's begin by checking the class of the plants variable with class(plants). This will give us a clue as to the overall
structure of the data.

```{r}
class(plants)
```

It's very common for data to be stored in a data frame. It is the default class for data read into R using functions

Since the dataset is stored in a data frame, we know it is rectangular. In other words, it has two dimensions (rows and
columns) and fits neatly into a table or spreadsheet. Use dim(plants) to see exactly how many rows and columns we're
dealing with.

```{r}
dim(plants)
```
The first number you see (5166) is the number of rows (observations) and the second number (10) is the number of
columns (variables).

You can also use nrow(plants) to see only the number of rows. Try it out.

```{r}
nrow(plants)
```
... And ncol(plants) to see only the number of columns.
```{r}
ncol(plants)
```
If you are curious as to how much space the dataset is occupying in memory, you can use object.size(plants).

```{r}
object.size(plants)
```
Now that we have a sense of the shape and size of the dataset, let's get a feel for what's inside. names(plants) will
return a character vector of column (i.e. variable) names. Give it a shot.

```{r}
names(plants)
```

We've applied fairly descriptive variable names to this dataset, but that won't always be the case. A logical next step
is to peek at the actual data. However, our dataset contains over 5000 observations (rows), so it's impractical to view
the whole thing all at once.

The head() function allows you to preview the top of the dataset. Give it a try with only one argument.

```{r}
head(plants)
```
Take a minute to look through and understand the output above. Each row is labeled with the observation number and each
column with the variable name. Your screen is probably not wide enough to view all 10 columns side-by-side, in which
case R displays as many columns as it can on each line before continuing on the next.
          
By default, head() shows you the first six rows of the data. You can alter this behavior by passing as a second
argument the number of rows you'd like to view. Use head() to preview the first 10 rows of plants.

```{r}
head(plants, 10)
```
The same applies for using tail() to preview the end of the dataset. Use tail() to view the last 15 rows.

```{r}
tail(plants, 15)
```
After previewing the top and bottom of the data, you probably noticed lots of NAs, which are R's placeholders for
missing values. Use summary(plants) to get a better feel for how each variable is distributed and how much of the
dataset is missing.

```{r}
summary(plants)
```

summary() provides different output for each variable, depending on its class. For numeric data such as Precip_Min,
summary() displays the minimum, 1st quartile, median, mean, 3rd quartile, and maximum. These values help us understand
how the data are distributed.
You can see that R truncated the summary for Active_Growth_Period by including a catch-all category called 'Other'.
Since it is a categorical/factor variable, we can see how many times each value actually occurs in the data with
table(plants$Active_Growth_Period).

```{r}
table(plants$Active_Growth_Period)
```

Each of the functions we've introduced so far has its place in helping you to better understand the structure of your
data. However, we've left the best for last....

Perhaps the most useful and concise function for understanding the *str*ucture of your data is str(). Give it a try
now.

```{r}
str(plants)
```
The beauty of str() is that it combines many of the features of the other functions you've already seen, all in a
concise and readable format. At the very top, it tells us that the class of plants is 'data.frame' and that it has 5166
observations and 10 variables. It then gives us the name and class of each variable, as well as a preview of its
contents.

str() is actually a very general function that you can use on most objects in R. Any time you want to understand the

In this lesson, you learned how to get a feel for the structure and contents of a new dataset using a collection of
simple and useful functions. Taking the time to do this upfront can save you time and frustration later on in your
analysis.

