---
title: "dplyr"
author: "Pradip Muhuri"
date: "7/15/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Manipulating Data with dplyr Package
https://www.geek-share.com/detail/2669229281.html#grouping-and-chaining-with-dplyr-package

dplyr is a fast and powerful R package written by Hadley Wickham and Romain Francois. The dplyr philosophy is to have small functions that each do one thing well.

One unique aspect of dplyr is that the same set of tools allow you to work with tabular data from a cariety of sources, including

* data frame

* data tables

* databases

* multidimensional arrays

Task 1: load the data into R.


```{r}
library(ISwR)
```
Task 2: data structure.
```{r}
str(mtcars)
```
Task 3: library dplyr package.

```{r}
library(dplyr)  
# check version: you need to have version 0.4.0 or later  
packageVersion("dplyr")  
```
Task 4: Find the class of mtcars.

```{r}
class(mtcars)
```

Task 5: Dimension of mtcars

```{r}
dim(mtcars)
```

Let us consider the five manipulation tasks with dplyr

- `select()`  
- `filter()`  
- `arrange()`  
- `mutate()`  
- `summarize()`  

Task 6: select three variables of stroke_tbl. There is no need to use the $ when specifying the variable names in dplyr.

```{r}
select(mtcars, mpg, wt, vs)
```

Task 7: select a sequence of columns.


```{r}
select(mtcars, mpg:drat)
```
Task 8: throw away selected columns.(the negative sign in front of drat tells us we don’t want those columns.)


```{r}
select(mtcars, -drat)

```
Task 9: throw away multiple columns.


```{r}
select(mtcars, -(drat:carb))
```
Task 10: use filter function to select all rows for which the cyl is equal to 6.

```{r}
filter(mtcars, cyl == 6)
```
Task 11: filter rows based on certain conditions. Those conditions (variables) are separated by commas.

```{r}
filter(mtcars, cyl ==  6, hp > 120)

```
https://www.datasciencemadesimple.com/filter-subsetting-rows-r-using-dplyr/


Task 12: subset mtcars rows with multiple conditions using %in% operator

```{r}
filter(mtcars, gear %in% c(3,5))
```
```{r}
filter(mtcars, gear %in% c(4,5) & carb==2)
```
Task 13: Filter the rows in R with multiple conditions (NOT). using Dplyr

```{r}
filter(mtcars, !gear %in% c(4,5))
```
Task 14: Filter the rows in R with Contains condition using Dplyr

```{r}
filter(mtcars, grepl(0,hp))
```
Task 15: Subset using Slice Family of function in R dplyr
 

```{r}
mtcars %>% slice_head(n = 5)
```
Task 16: Subset using Slice Family of function in R dplyr 
 

```{r}
mtcars %>% slice_tail(n = 5)
`````
Task 17: slice_max() function returns the max n rows of the dataframe based on a column as shown below.
```{r}
mtcars %>% slice_max(mpg, n = 5)
```
Task 18: slice_min() function returns the minimum n rows of the dataframe based on a column as shown below.
```{r}
mtcars %>% slice_min(mpg, n = 5)
```
Task 19: slice_sample() function returns the sample n rows of the dataframe as shown below.


```{r}
mtcars %>% slice_sample(n = 5)
```
Task 20: slice_head() by group in R:  returns the top n rows of the group using slice_head() and group_by() functions

```{r}
mtcars %>% group_by(vs) %>% slice_head(n = 2)

```
```{r}
mtcars %>% group_by(vs)
```
```{r}
sample_n(mtcars,4)
```