---
title: "with() vs. within()"
author: "Pradip Muhuri"
date: "7/16/2022"
output: pdf_document
---

Acknowledgements: The text and code are from https://www.journaldev.com/47815/with-and-within-function-in-r

The with() provides the output separately for every value associated with the columns of the data frame.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd('c:/r-basics/ManipulatingData')

```
```{r}
rm(list = ls())
 
Num <- c(100,100,100,100,100)
Cost <- c(1200,1300,1400,1500,1600)
 
data_A <- data.frame(Num,Cost,stringsAsFactors = FALSE)
 
with(data_A, Num*Cost)
with(data_A, Cost/Num)
```

The with() function does not alter the original data frame at any cost. 
```{r}
data_A
```

R within() function calculates the outcome of the expression within itself, but allowing us to create a copy of the data frame with a newcolumn that would eventually stores the result of the R expression.

```{r}
rm(list = ls())
 
Num <- c(100,100,100,100,100)
Cost <- c(1200,1300,1400,1500,1600)
data_A <- data.frame(Num,Cost,stringsAsFactors = FALSE)
 within(data_A, {Product <- Num*Cost; Q <- Cost/Num; data_A})
```
In order to save the addition of two columns create an extra column with $ symbol
```{r}
rm(list = ls())
 
Num <- c(100,100,100,100,100)
Cost <- c(1200,1300,1400,1500,1600)
data_A <- data.frame(Num,Cost,stringsAsFactors = FALSE)
data_A$Product <- Num*Cost
data_A$Q <- Cost/Num
data_A
```

